---
title: "Vectors and Descriptive Statistics"
subtitle: "R Syntax 3"
output: html_notebook
---

## Vectors: Collecting Values Together into a Single Variable Name

In R, it is rare for you to create dozens of variables; usually, putting
data together as a dataset into a single variable makes work simpler!

Data is grouped together into a data structure.  R supports many kinds of data structures (e.g. matrices, lists, dataframes, etc), but the simplest one is
the *vector*, a simple sequence of values.  We can get started on many statistics
just using vectors, and most operations on vectors work on the other data 
structures, making vectors a great place to get started in R.


## Creating Vectors

Many functions exist in R to quickly create vectors:

  - **`c(1, 3, 5)`**: combine values into the vector  *(1, 3, 5)*.
  - **`1:5`**: create a vector of the values 1 through 5 *(1, 2, 3, 4, 5)*
  - **`seq(1, 5)`**: Also creates a vector of 1 through 5 *(1, 2, 3, 4, 5)*
  - **`seq(1, 5, by=2)`**:  1 through 5, every 2 values  *(1, 3, 5)*
  - **`seq(1, 5, length.out=2)`** Two total values from 1 through 5 *(1, 5)*
  

**Exercises**

Make a vector "data" with the values 2, 2, 5, 7, 2
```{r}
data <- c(2, 2, 5, 7, 2)
data
```

Make a vector "data" with the values 4, 6, 7, 3, 4
```{r}

```


Make a vector "data" with the values 5, 3, 2
```{r}

```


Make a vector "data" with the values T, F, F, T
```{r}

```


Make a vector "data" with the values 2.6, 3.2, 1.1, 0.9
```{r}

```



Without using the c() function, make a vector "data" with the values 1, 2, 3, 4
```{r}

```


Without using the c() function, make a vector "data" with the values 1, 3, 5, 7
```{r}

```


Without using the c() function, make a vector "data" with the values -20 through 20
```{r}

```

Make a vector "data" with 10 values evenly space from 1 to 35
```{r}

```


Make a vector with the values (1, 2, 3, 4.4, 5).  What is strange about this
vector?
```{r}

```

Makea vector with the values (T, F, 3, F, T).  What is strange about this vector?
```{r}

```


## Summary Statistics on Vectors

Many functions take sequences of data like vectors as an input.  They are really
useful, because they don't care how much data you have!

Let's start with these functions: `sort()`, `length()` `sum()`, `mean()`.


**Exercises**

Sort this dataset in ascending order (i.e. from low to high): (6, 2, 5, 3, 4)
```{r}

```

Using an optional input to the `sort()` function:
Sort this dataset in descending order (i.e. from high to low): (6, 2, 5, 3, 4)
```{r}

```



Calculate the length of the vector (1, 4, 5).  What does the length function do?
```{r}

```

Calculate the sum of all the values from -5 to 5.
```{r}

```

Calculate the sum of all the values from one to one million!  Does it take long?
```{r}

```

Calculate the mean of the dataset (1, 5, 3, 2).
```{r}

```


Without using the `mean()` function, calculate the mean of the dataset (1, 5, 3, 2)
*Tip*: The mean is the sum of all the values divided by the total number of values.
```{r}

```


## Vectorized Tranformations on Vectors: R's Cycle Rule

All of the math and logical operations in R (in fact, most operatoins in R)
work as "vectorized" operations.  This means that if a calculation is meant for
one value, it is automatically repeated for every value in the data structure.

For example, to add 1 to all values from 10 to 15 is as simple as `10:15 + 1`
Let's try it out!


**Exercises**

Multiply all the values from 1 to 5 by 10.
```{r}
1:5 * 10
```


Subtract 1000 from all the values from 1 to 10.
```{r}

```


Check each value from 1 to 10 if it is greater than 5 (tip: answer should have 5 Fs and 5 Ts)
```{r}

```


Check each value from 1 to 10 if the value's square (power of 2) is less than 10.
```{r}

```


calculate the square root of all the values from 1 to 10.
```{r}

```


Calculate the sum of squares of the values 1 through 10

```{r}

```


Calculate the square root of the sum of squares of the values 1 through 10

```{r}

```



When math operations like the sum or mean are done on logical values (T and F),
R automatically converts them to 1 and 0, respectively.  This is a useful feature,
and helps to calculate **counts** and **proportions**.  Let's try it out!

How many values are there greater than 5 in this dataset: (1, 6, 3, 5, 7)?
```{r}
data <- c(1, 6, 3, 5, 7)
sum(data > 5)
```

What proportion (i.e. percentage) of values in the dataset are greater than 5?
(1 means 100%, and 0 is 0%)
```{r}
data <- c(1, 6, 3, 5, 7)
mean(data > 5)
```

How many values in the dataset (5, 7, -2, -4, -3) are negative?
```{r}

```


What proportion of values in the dataset 1 through 50 are greater than the
dataset's mean?
```{r}

```


How many values in the dataset (5, 8, 4, 6, 5, 7, 5) are equal to 5?
```{r}

```


What proportion of the values from 1 to 100 are divisible by 2 
(i.e. has a remainder of 0 after dividing the value by 2)?
(reminder: the `%%` operator is used to get the remainder from integer division).
```{r}

```


## Cycle Rule: Vectorizing along multiple vectors

This vectorization rule also applies when working with multiple vectors, letting
you add up vectors together.  The way R does this is called the **"Cycle Rule"**;
it cycles through each value in the vectors, doing the operation on each in turn.
It continues cycling through values until the *longest* vector is finished.


Add up all the values in dataset "data1" with the values 1 through 5
and all the values in dataset "data2" with the values 21 through 25
```{r}
1:5 + 21:25
```

Same as above, but creating variables first.
```{r}
data1 <- 1:5
data2 <- 21:25
data1 + data2
```


Add up all the values between the values 1 through 5 and 10 through 14.
```{r}

```


Multiply these two datasets together: (-1, 0, 1, 0, -1) and (1, 2, 3, 4, 5)
```{r}

```



## Subsetting Data: Selecting Values from Vectors

Once you've collected data together, you're not stuck with all of your data;
a subset of a dataset can also be extracted out!



```{r}
a <- 11:20
a[2:5]
?which
a[a > 15]
?subset
subset(a, a > 15)
a[which(a > 15)]
a[2:length(a)]
```







## Functions with a variable number of inputs


```{r}
av <- function(...){mean(c(...))}
av(3, 4, 5)
av(5)
av(6, 7, 8, 9)
```




```{r}
a <- function(..., x=5, y=3){x + y}

```

